---
title: "Positions most important skills using Fifa19 data"
author: "Mohamed Atef"
date: "May 6, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
```

# Introduction  
Football players have many different skills and each position require better skills than other positions. 
For example, the important skills that a team would need in a CB are, for sure, totally different of those required in a ST.  
In this markdown, I want to analyze Fifa19 dataset to find out a robust method that relates player skills to his most suitable position to play.

***
# Reading Data
```{r}
fifa19 <- read.csv("data/main_players.csv", stringsAsFactors = F)
```

Let's See the colnames of the dataset
```{r}
colnames(fifa19)
```

We need to define all possible positions and skills columns in the dataset.
```{r}
positions_names <- c("LS", "ST", "RS", "LW", "LF", "CF", "RF", "RW", "LAM", "CAM", "RAM", "LM", "LCM", "CM", "RCM", "RM", "LWB", "LDM", "CDM", "RDM", "RWB", "LB", "LCB", "CB", "RCB", "RB")
skills_names <- c("Crossing", "Finishing", "HeadingAccuracy", "ShortPassing", "Volleys", "Dribbling", "Curve", "FKAccuracy", "LongPassing", "BallControl", "Acceleration", "SprintSpeed", "Agility", "Reactions", "Balance", "ShotPower", "Jumping", "Stamina", "Strength", "LongShots", "Aggression", "Interceptions", "Positioning", "Vision", "Penalties", "Composure", "Marking", "StandingTackle", "SlidingTackle", "GKDiving", "GKHandling", "GKKicking", "GKPositioning", "GKReflexes")
```

Also we need to explore the first observations
```{r}
head(fifa19)
```

***
# Data Cleansing
It seems that the positions ratings columns need some cleansing. I can't understand what the "+" and its following number means and I don't think it is so significant that I need to keep them. Let's get rid of them.
```{r}
for (position in positions_names) {
  fifa19[position] <- lapply(fifa19[position], function(rating){
    as.integer(substr(rating, 0, 2))
  })
}
```

***
# Data Exploration  
## Univariate Analysis
Now, let's explore the distribution of the players ratings
```{r}
fifa19 %>%
  ggplot(aes(x = Overall)) + geom_histogram(bins = 50)
```

A perfect normal distribution!! What about the players individual skills? 
```{r}
skills_names %>%
  lapply(function(skill){
    ggplot(fifa19, aes_string(x = skill)) + geom_histogram(bins = 50)
  })
```

Obviously, all skills are somehow normally distributed.
Only the GK-related skills are not, or actually they are mixtures of other normal distributions.  
I think this is becuase a small set of the players (Actual GKs) have good GK-related skills and the other have so small such skills.
In each set, these GK-skills will be normally distributed, so we find 2 normal distributions, one to the right about real GKs and one to the left about other players.  

***

# Modeling
## Problem Formulation  
As I stated at first, my goal is to find the best skills that each position would need.
Using current dataset, I think I can fulfill this in the following way:  
  1. Filter Players of each position.  
  2. Fit a Linear model between the **Overall** rating and the **skills** ratings.  
My idea is that I think *EA Sports* has a specific linear combination, for each position, between the skills of the player that produces his overall rating. The idea of fitting a line can also be defended by the fact that all of the factors and target variables are normally distributed as shown previously.  

## Testing Criteria  
To avoid **HARKing** (Hypothesizing After the Results are Known), I would like to state a testing criteria to my method.  
If my idea is correct, then final models would satisfy the following:  
  - The linear model must be a linear combination of just the skills and has no intercept component.  
  - The R-squared of this model is so near to 1.  
  - For each position, the most important skills are logical. For example, for a CB, Strength or Interception is much more important than say Finishing.
  
After finding such a model, we can get the most important skills of each position by considering all the skills that have a p-value < minimum_accepted_threshold with the target variable. so now let's start  

## Grouping Positions  
I think there is no diference between the skills that EA Sports would consider for the players with the same roles on the two sides of the pitch.  
What I mean is that the skills of a RB, for example, is the same as those for a LB, so we would like to combine all these positions with shared skills in the same "grouping position".
```{r}
grouped_positions_names <- c("ST", "ST", "ST", "LW/RW", "CF", "CF", "CF", "LW/RW", "AM", "AM", "AM", "LM/RM", "CM", "CM", "CM", "LM/RM", "LB/RB", "DM", "DM", "DM", "LB/RB", "LB/RB", "CB", "CB", "CB", "LB/RB")
```

## Players' best positions
Now Let's try to find the best position for each player
```{r}
fifa19$Preferred.Position <- fifa19[positions_names]%>%
  apply(1, function(row){
    if(is.na(row["LAM"])){
      "GK"
    }else{
      grouped_positions_names[which.max(row)]
    }
  }) %>% unlist()
fifa19$Preferred.Position %>% table()
```

## Actual Modeling
At first we set the minimum_accepted_threshold. Say we would accept the skill if and only if it has a p-value < 0.01 to have a 99% confidence interval.
```{r}
minimum_accepted_threshold <- 0.01
```

Let's give it a try on CBs to test the idea and find if it is right.
```{r}
CB_model <- fifa19 %>%
  filter(Preferred.Position == "CB") %>%
  select(Overall, skills_names) %>%
  lm(formula = Overall ~ . - 1) %>% ### No interception
  summary()
```

What about the R-Squared metric?
```{r}
CB_model$r.squared
```
Woooooooooooooow!! 0.999856!! very close to 1!!It seems like it is working !!  
Now, Let's run the third test and explore the most important skills for a CB that the model suggests
```{r}
CB_model$coefficients %>% 
  as.data.frame() %>%
  rownames_to_column("Skill") %>%
  filter(`Pr(>|t|)` < minimum_accepted_threshold) %>%
  arrange(`Pr(>|t|)`)
```

As expected. The most important factors are as follows, ordered in a descending order by importance:
```{r, echo=FALSE}
CB_model$coefficients %>% 
  as.data.frame() %>%
  rownames_to_column("Skill") %>%
  filter(`Pr(>|t|)` < minimum_accepted_threshold) %>%
  arrange(`Pr(>|t|)`) %>%
  "$"(Skill)
```
So logical.  
The 3 conditions are satisfied!!  My idea seems to be working, so let's generalize it over all the positions we have.
```{r}
all_positions_models <- fifa19$Preferred.Position %>%
  unique() %>%
  lapply(function(position){
    fifa19 %>%
      filter(Preferred.Position == position) %>%
      select(Overall, skills_names) %>%
      lm(formula = Overall ~ . - 1) %>% ### No interception
      summary()
  })
names(all_positions_models) <- unique(fifa19$Preferred.Position)
```

Let's check our conditions. We built our models under the first condition, so we need only to test the second one which is about the R-Squared values that must be so close to 1.
```{r}
all_positions_models %>%
  lapply(function(model) model$r.squared) %>%
  unlist() %>%
  summary()
```
As we see, the minimum R-Squred value is 0.9992 which is so close to 1.  

***
# Results
  
As our models conditions are satisfied, we can now use these models to interpret the most important skills for each position.  
At first, let's keep only the features that have p-values < minimum_accepted_threshold
```{r}
positions_most_important_features <- all_positions_models %>%
  lapply(function(model){
    model$coefficients %>% 
      as.data.frame() %>%
      rownames_to_column("Skill") %>%
      filter(`Pr(>|t|)` < minimum_accepted_threshold) %>%
      arrange(`Pr(>|t|)`)
  })
names(positions_most_important_features) <- names(all_positions_models)
```

Now Let's get the most important skills of each position.
```{r}
positions_most_important_features
```
As we can see, the results are so logical.  
If we look at the most two important skills for each position, we find the following:  

Position | Most Important Skill
------------- | -------------
GK | Reactions, GKPositioning
CB | Strength, Marking
LB/RB | Stamina, Crossing
DM | Interceptions, Stamina
CM | ShortPassing, Stamina
AM | Vision, ShortPassing
LM/RM | Crossing, Dribbling
LW/RW | Dribbling, Crossing
ST | Finishing, Positioning

Results are so meaningful and it seems that our reverse feature engineering is fruitful!!  
This way I could reach my goal and find out the best features that characterise each position. Next time when someone need to scout a good player, he may use these results with no fear ;)  